Parsing Postfix log files

### 2009-03-26 doc/logparser-complications.tex
Dealt with some XXX items.

    \item If one or more recipients have been accepted, Postfix will have
        allocated a queueid for the incoming mail, and there may be a mail
        in the state tables that needs to be discarded.  The timeout may
        have interrupted transfer of an accepted delivery attempt, or it
        may have occurred after a mail delivery attempt was rejected.

        XXX THE REMAINDER OF THIS DOES NOT MAKE SENSE\@; CHECK AND CORRECT
        IT\@.  To distinguish between the two cases the action compares the
        timestamp of the last accepted mail against the timestamp of the
        last log line from \daemon{smtpd} for that connection; this depends
        on the \action{CLONE} action keeping a list of all mails accepted
        for each connection.  If the mail acceptance timestamp is later
        than the timeout timestamp, the timeout applies to the
        just-accepted mail, which will be removed from the state tables,
        without entering it in the database.  If the \daemon{smtpd}
        timestamp is later there was a rejection between the accepted mail
        and the timeout: the action assumes that the timeout applies to the
        rejected delivery attempt and finishes.  This assumption is not
        necessarily correct, because Postfix may have accepted an earlier
        recipient and rejected a later one, in which case the timeout
        applies to the accepted mail, which should be discarded; this has
        not been a problem in practice.  Processing timeouts is further
        complicated by the presence of out of order \daemon{cleanup} log
        lines: see \sectionref{discarding cleanup log lines} for details.

@@

\newpage{} %XXX VARIOREF PROBLEM

>>>>

    \item If one or more recipients have been accepted, Postfix will have
        allocated a queueid for the incoming mail, and there may be a mail
        in the state tables that needs to be discarded.  The timeout may
        have interrupted transfer of an accepted delivery attempt, or it
        may have occurred after a mail delivery attempt was rejected.  If a
        mail needs to be discarded, the following checks will all pass:

            \item The timestamp of the log line preceding the timeout log
                line is earlier than the timestamp of the last accepted
                delivery attempt, i.e.\ there have not been any rejections
                since then the delivery attempt was accepted.

            \item The mail must exist in the state tables.

            \item The mail does not have a \daemon{qmgr} log line.

        If all the checks are passed the mail will be discarded from the
        state table, and will not be entered in the database.  If one or
        more checks are not passed it is assumed that the timeout happened
        after a rejected delivery attempt.  This assumption is not
        necessarily correct, because Postfix may have accepted an earlier
        recipient and rejected a later one, in which case the timeout
        applies to the partially accepted mail, which should be discarded;
        this has not been a problem in practice.  Processing timeouts is
        further complicated by the presence of out of order
        \daemon{cleanup} log lines: see \sectionref{discarding cleanup log
        lines} for details.

@@

### 2009-03-26 doc/logparser-implementation.tex
Dealt with some XXX items.

Everything starts off with a mail entering the system, whether by local
submission via \daemon{postdrop}, by \acronym{SMTP}, by re-injection due to
forwarding, or internally generated by Postfix.  Local submission is the
simplest case: a queueid is assigned immediately \flowchart{PICKUP}{2}, and
the mail moves on to the delivery stage.  Re-injection due to forwarding
sadly lacks explicit log lines of its own; it is explained fully in
\sectionref{Re-injected mails}.  Internally generated mails lack any
explicit origin in Postfix 2.2.x and must be detected using heuristics as
described in \sectionref{identifying bounce notifications}; later versions
of Postfix do provide log lines for internally generated mails.  Bounce
notifications are the primary example of internally generated mails, though
there are other types, e.g.\ Postfix may generate mails to the
administrator when it encounters configuration errors, but such mails are
presumably rare.  XXX WOULD THIS PARAGRAPH BE BETTER AS A LIST\@?

@@

    \item This is followed by rejection of sender address, recipient
        addresses, client \acronym{IP} address or hostname, HELO hostname,
        etc.\ \flowchart{DELIVERY\_REJECTED}{4}; acceptance of one or more
        mails \flowchart{CLONE}{5}; or some interleaving of both.

@@

    \item [smtp\_code] The \acronym{SMTP} code associated with the log
        line.  In general an \acronym{SMTP} code is only present in a log
        line representing a rejection or final delivery; results whose log
        line did not contain an \acronym{SMTP} code will duplicate the
        \acronym{SMTP} code of other results in that connection.  Some
        final delivery log lines do not contain an \acronym{SMTP} code: in
        those cases the \acronym{SMTP} code is specified by the rule's
        \texttt{result\_data} field, based on the success or failure
        represented by the log line.  XXX GIVE AN EXAMPLE LOG LINE\@.

>>>>

Everything starts off with a mail entering the system, whether by local
submission via \daemon{postdrop}, by \acronym{SMTP}, by re-injection due to
forwarding, or generated internally by Postfix.  Local submission is the
simplest case: a queueid is assigned immediately \flowchart{PICKUP}{2}, and
the mail moves on to the delivery stage.  Re-injection due to forwarding
sadly lacks explicit log lines of its own; it is explained fully in
\sectionref{Re-injected mails}.  Internally generated mails lack any
explicit origin in Postfix 2.2.x and must be detected using heuristics as
described in \sectionref{identifying bounce notifications}; later versions
of Postfix do provide log lines for internally generated mails.  Bounce
notifications are the primary example of internally generated mails, though
there are other types, e.g.\ Postfix may generate mails to the
administrator when it encounters configuration errors, but such mails are
presumably rare.

@@

    \item This is followed by rejection of sender address, recipient
        addresses, client \acronym{IP} address or hostname, etc.\
        \flowchart{DELIVERY\_REJECTED}{4}; acceptance of one or more mails
        \flowchart{CLONE}{5}; or some interleaving of both.

@@

    \item [smtp\_code] The \acronym{SMTP} code associated with the log
        line.  In general an \acronym{SMTP} code is only present in a log
        line representing a rejection or final delivery; results whose log
        line did not contain an \acronym{SMTP} code will duplicate the
        \acronym{SMTP} code of other results in that connection.  Some
        final delivery log lines do not contain an \acronym{SMTP} code,
        e.g.\ when Postfix delivers to a command; in those cases the
        \acronym{SMTP} code is specified by the rule's
        \texttt{result\_data} field, based on the success or failure
        represented by the log line.
